---
title: "README"
author: "Bob Rudis"
date: January 11, 2015
output:
  md_document:
    variant: markdown_github
---

The CDC's FluView is a Flash portal and the only way to get flu season
data is to use GUI controls, making it tedious to retrieve updates. This package
uses the same API the portal does to programmatically retrieve data.

The following functions are implemented:

- `get_flu_data` : retrieve flu data (national, by various region/sub-region types)
- `get_state_data` : retrieve state-level flu data

The following data sets are included:

### News

- Version 0.1 released
- Version 0.2 releases : added state-level data retrieval

### Installation

```{r eval=FALSE}
devtools::install_github("hrbrmstr/cdcfluview")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r}
library(cdcfluview)
library(ggplot2)
library(dplyr)
library(statebins)

# current verison
packageVersion("cdcfluview")

flu <- get_flu_data("hhs", sub_region=1:10, "ilinet", years=2014)

state_flu <- get_state_data()

dplyr::glimpse(flu)

gg <- ggplot(flu, aes(x=WEEK, y=X..WEIGHTED.ILI, group=REGION))
gg <- gg + geom_line()
gg <- gg + facet_wrap(~REGION, ncol=2)
gg <- gg + theme_bw()
```

```{r echo=FALSE, fig.height=10, fig.width=6}
gg
```

```{r}
gg_s <- state_flu %>%
  filter(WEEKEND=="Jan-03-2015") %>%
  select(state=STATENAME, value=ACTIVITY.LEVEL) %>%
  filter(!(state %in% c("Puerto Rico", "New York City"))) %>% # need to add PR to statebins
  mutate(value=as.numeric(gsub("Level ", "", value))) %>%
  statebins(brewer_pal="RdPu", breaks=4, 
            labels=c("Minimal", "Low", "Moderate", "High"),
            legend_position="bottom", legend_title="ILI Activity Level") +
  ggtitle("CDC State FluView (2015-01-03)")
```

```{r echo=FALSE, fig.height=5, fig.width=7}
gg_s
```

### Test Results

```{r}
library(cdcfluview)
library(testthat)

date()

test_dir("tests/")
```

